## Mapping

We'll need a few new packages to make this work: the [leaflet](https://rstudio.github.io/leaflet/) package for interactive maps, and the [viridis](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html) package for adding color!

```{r echo = FALSE}
data_geo <- read.csv("data/data_geocoded.csv")
library(tidyverse)
```


```{r warning = FALSE, message = FALSE}
library(leaflet)
library(viridis)
```

I'm most interested in the prices of the campsites, so I'll categorize them into four groups: "free", "cheap" (less than $12/night, as defined by [freecampsites.net](https://freecampsites.net/about/)), "moderate" (about double "cheap"), and "expensive" (as far as sleeping in a tent is concerned).

```{r}
data_geo <- data_geo %>% 
    mutate(price_cat = case_when(
        price == 0 ~ "free",
        price > 0 & price <= 12 ~ "cheap",
        price > 12 & price <= 25 ~ "moderate",
        price > 25 ~ "expensive"))
```

Along with the categories, I need to define a color palette. Here, I've just used the viridis pre-sets.

```{r}
pal <- colorFactor(viridis_pal()(4), 
                   domain = c("free", "cheap", "moderate", "expensive"))

```

With just a few lines of code, my map is ready to go. Voila!

```{r message = FALSE, warning = FALSE}
m <- leaflet(data_geo) %>%
    addTiles() %>%  
    addCircleMarkers(color = ~pal(price_cat),
                     popup = paste(data_geo$campground,
                                   data_geo$price,
                                   sep = ", $"))
m 
```

<br><br>
In green are the free sites, purple is "cheap", yellow is "moderate", and blue is "expensive". If you click on a site, you can see the campground (or sometimes parking lot/airbnb) name and the price.